How to publish in Biological Sciences: a guide for the uninitiated

### 2021-12-15 2.07-submit.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

## A typical submmission work-flow

The Parts A to G (below) all refer back to letters in circles in Figure \@ref(fig:Editorial-work-flow). Remember that most of the actions described below are also the subject of later sections in [Parts 2](#part2) [and 3](#part3) of this book.

### Part A - formatting and submitting

@@

### Part B - Editorial assistant and/or editor check the manuscript

@@

### Part C - Associate Editor takes over

 - The AE [should read the manuscript](#editors) and may [desk reject](#desk) it if they feel that it won’t make it through review. An AE [rejection](#rejection) isn’t great news as AEs don’t always have the best experience in knowing what will and what won’t make it through the review process. The editor usually has more experience. Hopefully, this won’t have taken long ([1-2 weeks](#timetaken)) and so won’t be very painful. This rejection may be fair or unfair, but it’s done and there’s nothing much you can do except return to Part A (above), but see [later sections](#rejection).
 - Normally, your manuscript will be sent out to review and you can expect to wait 4-6 weeks (good), but sometimes [up to 3 months](#timetaken), for a decision. If it’s away for over 3 months, you should definitely make a query on the editorial management system. Unsurprisingly, authors feel more unhappy with the review results the longer the process takes [@jiang2021understanding].

### Part D - Reading the reviews

 - Once back from review, you’ll get an email from the editorial management system with the decision.
 - __Reject and resubmit:__ This is a category that means you need major revision, but the journal doesn’t want the time that it takes to do this on their journal processing statistics. In many journals, this result has replaced ‘major revision’. Back to step 3 with a track changes manuscript and response letter to reviewer comments.
 - __Major revision:__ essentially the same as 6b. Both 6b and 6c result in your manuscript being reviewed again. You’ll need to carefully prepare both the manuscript and the response to reviewers as the reviewers will see both. Back to step 3 with a response letter.
 - __Minor revision:__ is unusual at this stage, but your manuscript should now only be assessed by the AE, so you should address your responses to them.
 - __Accept without revision:__ is practically unheard of. Possibly, your manuscript has already undergone some peer review (maybe as a preprint or in another journal). However, if you get this result after the first round of peer review, it's time to eat cake.

### Part E - Revising and resubmitting your manuscript

 - If you are resubmitting, aim to prioritise this to get it done in 2 to 3 weeks if possible. 
  + The reason is that the same reviewers are likely to be willing to look at your manuscript again within a month, and will remember all the points that they made. Similarly, the AE will remember all of the issues that they had. It’s hard to stress how valuable this is, as keeping it all fresh will result in a swift response.
  + If you don’t or can’t manage to get your responses back quickly, you might expect a rocky ride through the review process when you go back for the second round. The reviewers you had before might not be available, but the AE will be obliged to have at least 2 reviews again. This means that you may get new reviews. New reviewers are likely to throw up new issues, and could result in your manuscript getting rejected at this stage, or that you’ll have another major review decision, sending you back to step 3 with a track changes manuscript and response letter to reviewer comments. This drags the whole process on for much longer and reviewers and AEs are likely to look less favourably at your manuscript. 
  - A better result is when there are only minor revisions. In this case the manuscript is simply bouncing between you and the AE and even if this happens more than once, it’s fine as long as you can keep the response time reasonable (within a couple of weeks).
  - In either case, your [rebuttal](#rebuttal) letter will be a very important part of your resubmission, and this will be covered in a [later section](#rebuttal).

### Part F - Associate Editor recommends to Editor-in-Chief

@@

### Part G - acceptance

- Hopefully your manuscript will now be accepted, and you are entering the last stages of the process. Your accepted manuscript should be sent to the publishers for typesetting, and you can get the proofs back very quickly (for some publishers). Most demand that the proofs are returned very quickly (often within 48 hours), and you should try to prioritise this if you can. If you can, please also send it to the co-authors. The more eyes the better at this stage for spotting errors. Don’t expect to be able to change a lot in the proof process, it’s really just for catching errors. Carefully check all figures, tables and legends. It’s not unknown that typesetters cause problems when they make proofs (tables can be disasters). A detailed section is provided in [Part 3](#accepted) on what to do once your manuscript is accepted.
- If you (or a co-author) spot a fundamental error with your data or analyses at this point (or any of the other steps above), you should discuss it with all co-authors and decide what to do. It’s better to withdraw the manuscript now than to have to retract it later ([see part 4](#retract)).

### Rejected

- If your manuscript is [Rejected](#rejection), take the comments of the reviewers on board. Think about it for a couple of days, and then set about revising the manuscript. However unfair you think the reviewers have been, there should be some important messages for you to consider carefully and discuss among the co-authors before going back to step 2 with the next journal on your list. More information to reflect on regarding a rejection is provided in a [later section](#rejection).

## Remember that peer review is conducted by humans

The peer review process is not ideal, but it is worth remembering that it’s there to help improve your manuscript. The most prominent problems involve the [time](#timetaken) that the editorial team take to find reviewers and have them agree to complete reviews in a [timely manor](#timetaken). In Figure \@ref(fig:Editorial-work-flow), only four of the arrows are in the control of the authors, while at least 16 are within the editorial management process. Each one could be the source of your manuscript getting stuck. Throughout the process, you should be able to track the progress of your manuscript using the editorial management software online. More information on delays in the handling of manuscripts is provided in a [later section](#timetaken).

>>>>

## A typical submission work-flow

The steps A to G (below) all refer back to letters in circles in Figure \@ref(fig:Editorial-work-flow). Remember that most of the actions described below are also the subject of later sections in [Parts II](#part2) [and III](#part3) of this book.

**A - formatting and submitting**

@@

**B - Editorial assistant and/or editor check the manuscript**

@@

**C - Associate Editor takes over**

 - The AE [should read the manuscript](#editors) and may [desk reject](#desk) it if they feel that it won’t make it through review. An AE [rejection](#rejection) isn’t great news as AEs don’t always have the best experience in knowing what will and what won’t make it through the review process. The editor usually has more experience. Hopefully, this won’t have taken long ([1 to 2 weeks](#timetaken)) and so won’t be very painful. This rejection may be fair or unfair, but it’s done and there’s nothing much you can do except return to  **A** (above), but see [later sections](#rejection).
 - Normally, your manuscript will be sent out to review and you can expect to wait 4 to 6 weeks (good), but sometimes [up to 3 months](#timetaken), for a decision. If it’s away for over 3 months, you should definitely make a query on the editorial management system. Unsurprisingly, authors feel more unhappy with the review results the longer the process takes [@jiang2021understanding].

**D - Reading the reviews**

 - Once back from review, you’ll get an email from the editorial management system with the decision.
 - __Reject and resubmit:__ This is a category that means you need major revision, but the journal doesn’t want the time that it takes to do this on their journal processing statistics. In many journals, this result has replaced ‘major revision’. Back to step 3 with a track changes manuscript and response letter to reviewer comments.
 - __Major revision:__ essentially the same as reject and resubmit. Both reject and resubmit and Major revision result in your manuscript being reviewed again. You’ll need to carefully prepare both the manuscript and the response to reviewers as the reviewers will see both. Back to step C with a response letter.
 - __Minor revision:__ is unusual after the first round of review, so if you get this decision first time it's something to celebrate. You may have already done one (or more) rounds of Major revision before you get here. After addressing the Minor revisions, your manuscript should now only be assessed by the AE, so you should address your responses to them.
 - __Accept without further revision:__ is practically unheard of on first submission. Your manuscript may have already undergone some peer review (maybe as a preprint or in another journal), or in previous rounds of review. This decision is likely a recommendation from the AE to the Editor-in-Chief.

**E - Revising and resubmitting your manuscript**

 - If you are resubmitting, aim to prioritise this to get it done in 2 to 3 weeks if possible. 
  + The reason is that the same reviewers are likely to be willing to look at your manuscript again within a month, and will remember all the points that they made. Similarly, the AE will remember all of the issues that they had. It’s hard to stress how valuable this is, as keeping it all fresh will result in a swift response.
  + If you don’t or can’t manage to get your responses back quickly, you might expect a rocky ride through the review process when you go back for the second round. The reviewers you had before might not be available, but the AE will be obliged to have at least 2 reviews again. This means that you may get new reviews. New reviewers are likely to throw up new issues, and could result in your manuscript getting rejected at this stage, or that you’ll have another Major revision decision, sending you back to step C with a track changes manuscript and response letter to reviewer comments. This drags the whole process on for much longer and reviewers and AEs are likely to look less favourably at your manuscript. 
  - A better result is when there are only Minor revisions. In this case the manuscript is simply bouncing between you and the AE and even if this happens more than once, it’s fine as long as you can keep the response time reasonable (within a couple of weeks).
  - In either case, your [rebuttal](#rebuttal) letter will be a very important part of your resubmission, and this will be covered in a [later section](#rebuttal).

**F - Associate Editor recommends to Editor-in-Chief**

@@

**G - acceptance**

- Hopefully your manuscript will now be accepted, and you are entering the last stages of the process. Your accepted manuscript should be sent to the publishers for copy-editing and typesetting, and you can get the proofs back very quickly (for some publishers). Most demand that the proofs are returned very quickly (often within 48 hours), and you should try to prioritise this if you can. If you can, please also send the proofs to the co-authors. The more eyes the better at this stage for spotting errors. Don’t expect to be able to change a lot in the proof process, it’s really just for catching errors. Carefully check all figures, tables and legends. It’s not unknown that typesetters cause problems when they make proofs (tables can be disasters). A detailed section is provided in [Part III](#accepted) on what to do once your manuscript is accepted.
- If you (or a co-author) spot a fundamental error with your data or analyses at this point (or any of the other steps above), you should discuss it with all co-authors and decide what to do. It’s better to withdraw the manuscript now than to have to retract it later ([see part IV](#retract)).

**Rejected**

- If your manuscript is [Rejected](#rejection), take the comments of the reviewers on board. Think about it for a couple of days, and then set about revising the manuscript. However unfair you think the reviewers have been, there should be some important messages for you to consider carefully and discuss among the co-authors before going back to step A with the next journal on your list. More information to reflect on regarding a rejection is provided in a [later section](#rejection).

### Remember that peer review is conducted by humans

The peer review process is not ideal, but it is worth remembering that it’s there to help improve your manuscript. The most prominent problems involve the [time](#timetaken) that the editorial team take to find reviewers and have them agree to complete reviews in a [timely manner](#timetaken). In Figure \@ref(fig:Editorial-work-flow), only 4 of the arrows are in the control of the authors, while at least 16 are within the editorial management process. Each one could be the source of your manuscript getting stuck. Throughout the process, you should be able to track the progress of your manuscript using the editorial management software online. More information on delays in the handling of manuscripts is provided in a [later section](#timetaken).

### 2021-12-15 2.08-peerreview.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

## What are peer reviewers asked to do?
The peer review report consists of three major parts: the review, confidential comments to the editor, and the reviewer's opinion of what decision should be made. You will only have access to the first of these three. Understanding peer review is greatly aided by conducting peer review, which is covered by a section in [Part 4](#howpeerreview). Reading about [how to conduct peer review](#howpeerreview) may help with your interpretation of the reviews that you receive.

@@

The reviewer should then provide a general critique including positive and any negative aspects of the manuscript. They should provide detailed information on exactly how the manuscript should be improved, including any significant literature that might be missing from the manuscript. Lastly, they can provide a list of minor comments along the length of the manuscript that require further attention from the authors. When I undertake this last part, I tend to do it with page and line numbers (which is one reason why submitting a manuscript with line numbers is so important. If the minor comments get too numerous, then I tend to stick with major comments.

@@

### Reviewer's opinion of what the editorial decision should be
In many journals, the reviews are directly asked whether the manuscript should be rejected and resubmitted major or minor revisions. Personally I don't think that reviewers should be asked these questions as this is a decision for the editor after having read the reviewers comments and the manuscript.

@@

### The myth of the 'shit sandwich' {#shit}
The shit sandwich is where a review where the beginning and end are generally positive, while the very critical appraisal happens in the middle. Although the shit sandwich might be seen as a way for a reviewer to sugar coat their negative message, an analysis of PLOS ONE reviews by Eve et al. [-@eve2021reading] suggests that some of the best reviews (think Minor Revisions) actually have this format. Here the reviewer will be positive in the outset and the summing up, but then in the middle have a set of issues that need correcting. Hence, they found that the shit in the shit sandwich wasn't that bad. In the same analysis, Eve et al. [-@eve2021reading] found that truly bad (i.e. Reject) reviews could be bad at any point in their length.

@@

### Peer review is not a trial by committee
Those of you who have experienced manuscripts being critiqued at a book journal club will know that there are very few published papers that leave a journal club without having many negative, critical comments. Instead peer review is conducted by an individual, on their own and with their own personal limitations. Peer reviewers are forbidden from sharing the contents of a manuscript with others, without express permission from the editor.

### Correct English grammar
It's not the job of a reviewer to correct any faulty grammar on a manuscript. similarly it is not up to the reviewer to correct stylistic aspects of the manuscript. However, as English is such a subjective language, it is important that ambiguity is removed, and grammatical aspects can be important for this.

@@

### Tell them to cite other papers from the journal
This also happens [see section on Impact Factor](#impactfactor) often at the request of the editor.

@@

### Ignore what is good
Is often thought that peer review provides only negative criticism [see @eve2021reading]. This should not be the case as peer reviewers should also be able to accentuate the positive aspects of manuscripts that they read. Even if a manuscript is not considered publishable the positive attributes should be bought the attention of the editors as well as those that are negative.

>>>>

## What are peer reviewers asked to do?
The peer review report consists of three major parts: the review, confidential comments to the editor, and the reviewer's opinion of what decision should be made. You will only have access to the first of these three. Understanding peer review is greatly aided by conducting peer review, which is covered by a section in [Part IV](#howpeerreview). Reading about [how to conduct peer review](#howpeerreview) may help with your interpretation of the reviews that you receive.

@@

The reviewer should then provide a general critique including positive and any negative aspects of the manuscript. They should provide detailed information on exactly how the manuscript should be improved, including any significant literature that might be missing from the manuscript. Lastly, they can provide a list of minor comments along the length of the manuscript that require further attention from the authors. When I undertake this last part, I tend to do it with page and line numbers (which is one reason why submitting a manuscript with line numbers is so important). If the minor comments get too numerous, then I tend to stick with major comments.

@@

### Reviewer's opinion and the editorial decision {#AKP}
In many journals, the reviewers are directly asked whether the manuscript should be rejected or resubmitted (major or minor revisions). This opinion is given in a set of 'radio buttons' in the editorial management software. Depending on the journal, a number of additional qualitative questions are asked of reviewers that may be pivotal to the progress of a manuscript. These questions are set by the gatekeepers for the journal in question, and may ask whether the manuscript is worthy of publication in that journal, in the top 5% of important findings, or similar. Bornmann and Marx [-@bornmann2012anna] found that some journals require that all reviewers respond with a positive criterion on these questions, or the manuscript will not be considered further. They term this the 'Anna Karenina Principle' (AKP) of peer review because it requires 'positive selection' on all criteria, whereas previously papers were selected by meeting minimum standards; 'negative selection', which is now the overriding principle for '[no impact factor](#noIF)' journals.

Unless you know someone who can tell you the editorial practice of the journal you submitted to, you will not know whether or not your manuscript has been judged on the answers to these questions. However, if you receive all positive reviews and a rejection, it may well be that one or more of the positive reviewers did not select a criterion high enough on one of these questions. Personally, I don't think that reviewers should be asked these subjective questions, as this potentially trivialises decision making for gatekeepers. [Editors must read](#editors) the manuscript and all reviewer comments in order to reach decisions. Certainly, all journals need to provide [transparent criteria](#transparent) for selection of manuscripts, without which the black box nature via AKP workings of peer review will only perpetuate.

@@

### The myth of the 'shit sandwich' {#shit}
The shit sandwich is a review where the beginning and end are generally positive, while the very critical appraisal happens in the middle. Although the shit sandwich might be seen as a way for a reviewer to sugar coat their negative message, an analysis of *PLOS ONE* reviews by Eve et al. [-@eve2021reading] suggests that some of the best reviews (think Minor Revisions) actually have this format. Here the reviewer will be positive in the outset and the summing up, but then in the middle have a set of issues that need correcting. Hence, they found that the shit in the shit sandwich wasn't that bad. In the same analysis, Eve et al. [-@eve2021reading] found that truly bad (i.e. Reject) reviews could be bad at any point in their length.

@@

### Peer review is not a trial by committee
Those of you who have experienced manuscripts being critiqued at a journal club will know that there are very few published papers that leave a journal club without having many negative, critical comments. Instead peer review is conducted by an individual, on their own and with their own personal limitations. Peer reviewers are forbidden from sharing the contents of a manuscript with others, without express permission from the editor.

### Correct English grammar
It's not the job of a reviewer to correct any faulty grammar on a manuscript. Similarly it is not up to the reviewer to correct stylistic aspects of the manuscript. However, as English is such a subjective language, it is important that ambiguity is removed, and grammatical aspects can be important for this.

@@

### Tell them to cite other papers from the journal
This also happens [see section on Impact Factor](#increaseIF) often at the request of the editor. It is part of IF manipulation, and you should not act on this.

@@

### Ignore what is good
It is often thought that peer review provides only negative criticism [see @eve2021reading]. This should not be the case as peer reviewers should also be able to accentuate the positive aspects of manuscripts that they read. Even if a manuscript is not considered publishable the positive attributes should be bought the attention of the editors as well as those that are negative.

### 2021-12-15 2.10-editors.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

There is a worrying increase in poor editorial decision making, without any basis, because editors are not reading submissions. In their survey of Associate Editors, Poulson-Ellestad et al. [-@poulson-ellestad2020illuminating] found that they advised Early Career Researchers not to take on the position of Associate Editor unless they know that they have enough time in  their jobs. This is an assessment with which I agree, as you must be prepared to carve an hour or more of concentrated time several times a week, or nearly every day for the Editor-in-Chief.

When a manuscript is submitted to a journal, the submission goes to either the editor-in-chief or a handling editor based on the key words or journal section implied during submission (see [Part II](#submit)). In some journals (like [PeerJ](www.peerj.com)) the submissions are offered up to a whole group of editors who can take their pick. It seems that the next thing that happens is that the manuscript is sent out for peer review. But stop. That’s not correct and it’s really not a good way to proceed. Before sending it out, the designated handling editor needs to read the submission.

@@

## If you are going to be an editor, then you must be prepared to read
I must admit that I’ve done it. I’ve sent out manuscripts to be reviewed even when I didn’t have the time to properly read the article myself. A a superficial skim suggested that it seemed fine. Not good. It’s embarrassing to have sent out manuscripts that should be rejected without peer review. In the case I’m thinking of, once I’d found time to read through the manuscript later on that day, I realised how bad it was and immediately wrote to those I’d asked to do the reviews and asked them not to. The article was rejected. It is important that this burden is taken on by the editor than burden two or three times as many other reviewers to make the same call.

Sometimes, it’s not clear whether or not a manuscript will pass muster. Articles can stand or fall on good or bad single judgements of the authors. But misjudgements aren’t always obvious to editors. That’s why peer review is important, and that’s why it’s hugely important for editors to send manuscripts to appropriate reviewers that have some expertise in a subject.

@@

In my example above, the authors might assume that they had correctly identified the species when recording its call. Such assumptions should be backed up with museum and/or tissue bank accessions. But when they are not, the assumption that the authors are recording what they think they are, is vital. If this is placed in doubt, then the entire premise (description of a call to distinguish this species in the field) simply falls apart. In a case without vouchers, the assumption needs backing up by someone who knows the identification from another study, or without any foundation it becomes worthless.

>>>>

There is a worrying increase in poor editorial decision making, because editors are not reading submissions. In their survey of Associate Editors, Poulson-Ellestad et al. [-@poulson-ellestad2020illuminating] found that they advised Early Career Researchers not to take on the position of Associate Editor unless they know that they have enough time in  their jobs. This is an assessment with which I agree. You must be prepared to carve out dedicated time in which you can concentrate several times a week, or nearly every day for the Editor-in-Chief.

When a manuscript is submitted to a journal, the submission goes to either the editor-in-chief or a handling editor based on the key words or journal section implied during submission (see [Part II](#submit)). In some journals (like [*PeerJ*](www.peerj.com)) the submissions are offered up to a whole group of editors who can take their pick. It seems that the next thing that happens is that the manuscript is sent out for peer review. But stop. That’s not correct and it’s really not a good way to proceed. Before sending it out, the designated handling editor needs to read the submission.

@@

## If you are going to be an editor, then you must be prepared to read
I must admit that I’ve done it. I’ve sent out manuscripts to be reviewed even when I didn’t have the time to properly read the article myself. A superficial skim suggested that it seemed fine. Not good. It’s embarrassing to have sent out manuscripts that should be rejected without peer review. In the case I’m thinking of, once I’d found time to read through the manuscript later on that day, I realised how bad it was and immediately wrote to those I’d asked to do the reviews and asked them not to. The article was rejected. It is important that this burden is taken on by the editor than burden two or three times as many other reviewers to make the same call.

Sometimes, it’s not clear whether or not a manuscript will pass muster. Articles can stand or fall on good or bad single judgements of the authors. But misjudgements aren’t always obvious to editors. That’s why peer review is important, and that’s why it’s hugely important for editors to send manuscripts to appropriate reviewers who have some expertise in a subject.

@@

### 2021-12-15 3.00-part3.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

This part of the book gives you some things that you might want to do after your paper is accepted for publication. Much of this chapter is concerned with publicising the results and content beyond academia and for non-academic audiences. Even if you feel that there would be no interest beyond your academic niche, it would still be worth making some effort to popularise your study.

@@

In this millennium, the VoR had changed from a hard copy that was printed and bound into an issue of a journal into a pdf that appears at a journal website online. In addition, the VoR no longer has to belong to a volume or an issue, and is usually the first version available online [@haustein2015when]. The VoR can appear online long before it appears in an issue or volume (i.e. without page numbers), but still be the VoR.

@@

 - This means that if you wanted to make any changes to this first printed version (the VoR), you'd need to publish a separate [corrigendum](#corrigendum). 
 - The date of the VoR has an impact on primacy
  + for example in taxonomy, if there are two descriptions of a species, the earliest one counts as the valid one.

>>>>

This part of the book gives you information on some actions that you might want to do after your paper is accepted for publication. Much of this section is concerned with publicising the results and content beyond academia and for non-academic audiences. Even if you feel that there would be no interest beyond your academic niche, it would still be worth making some effort to publicise and popularise your study.

@@

In this millennium, the VoR had changed from a hard copy, that was printed and bound into an issue of a journal, into a pdf that appears at a journal website online. In addition, the VoR no longer has to belong to a volume or an issue, and is usually the first version available online [@haustein2015when]. The VoR can appear online long before it appears in an issue or volume (i.e. without page numbers), but still be the VoR.

@@

 - This means that if you wanted to make any changes to this first printed version (the VoR), you'd need to publish a separate [corrigendum](#corrigendum)
 - The date of the VoR has an impact on [primacy](#whatjournalfor)
   + in taxonomy, for example, if there are two descriptions of a species, the earliest one counts as the valid one

### 2021-12-15 3.01-accepted.Rmd
Add files via upload
# with lots of changes following copy edits and proofs

The day your paper is accepted, tell your advisor and co-authors. If they are in the same physical location as you then buy a cake and celebrate with them at tea time. If cake isn’t your thing, then find another appropriate treat for you and your co-authors. It’s a great achievement and something you should share together. If you aren’t in the same location then make a plan for next time you meet together, even if that’s only online.

Remember to make sure that all of your coauthors have a copy of the accepted version of the manuscript. Sometimes referred to as a postprint, you and your co-authors should submit this to your institutional repository so that the article can be reached as [Green Open Access](openaccess.html) by anyone who is interested in reading it.

@@

# Take your time with proofs {#proofs}

The next step in the publishing process once your paper has been accepted is that it will go for type setting. Depending on the journal and the publisher this process can proceed in several ways. Typically you will receive a notification that your paper proofs already and that you need to check them within 48 hours.

If you haven't checked proofs before then it is important that you read the instructions from the publishers carefully. They should tell you exactly what to do and if you are unsure about anything then talk to your co-authors.

Typically the publishers will send you a set of queries that relate to your proofs. They always ask you to check every author's name and affiliation. Other typical errors are that there are citations in the text that are not in the references. Or that there's literature in the references that are not cited.

The process of checking the proofs is very important. Errors can creep in during the type setting stage. Pay special attention to the tables table legends and figure legends.

You may also have the opportunity to change the size or orientation of figures if it looks like they are not well presented in the proof. Especially if the journal prints into columns they may choose to put your figure in one column instead of two. Another option is to have your figure in landscape across the whole page. Journals are generally pushed for space and so may refuse some requests for more room for larger figures. But you might get lucky if you make a good case.

Although the proofs are the responsibility of the corresponding author, it's good to get as many eyes on them as possible in order to spot any possible errors. Some journals let you know when proofs are likely to arrive, in which case it's a good idea to alert your co-authors and ask them whether they are prepared to look at them. I usually suggest that you make all your own corrections first before sending them around.

Many publishers want to proofs back in a hurry (typically 48 hours). If you don't have the confidence to correct proofs yourself, or cannot pass it around your co-authors within this deadline, then you can simply write back and ask for an extended deadline for your proofs. It is important to get it right, and much better than having to correct the paper.

@@

Probably the easiest way of doing proofs is to print them out and go through them with a pencil first. This allows you to take your time and you're more likely to spot errors this way than on the screen. However, some publishers will require you to submit proofs in an online system (effectively working with their LaTex document). You should still have an opportunity to print and take your time with the proofs though.

Sometimes, there are a lot of problems with proofs, and you may not have the confidence that the publishers will make all the corrections as indicated. In this case, you should ask to see another round of proofs before committing.

@@

Sometimes your article may acquire several DOIs, for example from the publisher and from the preprint server. Sites like Figshare and ResearchGate, assign different DOIs to content uploaded to their platforms. These are DOIs but they’re assigned by DataCite, another DOI registration agency. The different registration agencies provide different services that relate to DOIs and their associated metadata (from construction to the movie industry), and have different requirements for their member organisations. To prevent ambiguity, you should always use the DOI associated with the [Version of Record](#VoR).

@@

Publishers normally deposit DOIs and other metadata (authors' names and addresses, publication dates, title, licence, funders, etc.) around the time the article is published online. This is called content registration. This is not just for articles, but also preprints, conference proceedings, books, book chapters, peer reviews and more can all be registered with [Crossref](https://www.crossref.org/). The standardisation of this metadata means that not only is it possible to immediately find your article (the DOI is a link), but all of the metadata associated with it can be cross-referenced. This is of great help to funders, for example, who can look up all the products that their funding has produced.

### DOI Tools
If you can't find the DOI for an article that you want to cite, then there is this very useful online software that will provide the DOI if there is one for every reference you enter: [doi.crossref.org](https://doi.crossref.org/simpleTextQuery) or use the Crossref Metadata Search: https://search.crossref.org/references

There’s an equally useful database that provides BibTex for DOIs that you enter: [doi2bib.org](https://doi2bib.org/)

## Add your paper to your ORCID account {#ORCID}
The [ORCID](https://orcid.org/) (https://orcid.org/) number was devised to provide a common platform for authors to curate, and now some journals and funding authorities won’t allow you to submit or apply without one. This helps when there are authors with common (or even identical) names, and as an NGO dedicated to helping scientists ascribe credit for their work, should be supported. ORCID will also help prevent author fraud (see [Part I](authors.html)).

If you added you ORCID number to the metadata when you uploaded your manuscript, your ORCID record should be updated automatically (through Crossref) when the DOI of your article appears online. Inside your ORCID account, you can grant permission for Crossref to update the records automatically, otherwise you will need to log onto ORCID from time to time and approve changes suggested there. Alternatively, you can add works to your ORCID record via [Crossref Metadata Search](https://search.crossref.org/references).

Be careful not to create duplicate accounts, and make your ORCID available to your collaborators so that they aren't left guessing at your identity when it comes to submitting a paper. ORCID allows you to create some useful links for your website or a QR code for adding to posters and presentations. If you find you have a duplicated account, it is simple to remove it (see [here](https://support.orcid.org/hc/en-us/articles/360006896634)).

## Once you have a publication date
Once your paper is published you have an opportunity to publicise it yourself. There are lots of different ways to do this, [see the next chapter](citations.html).

## On the day you publish
This is a great opportunity to contact all the people who helped you in your study and send them a PDF of your paper. The easiest way to do this is to go to the acknowledgements section and write an email that includes everyone that is mentioned in the acknowledgements. write them a nice email in which you thank them for their help and explain briefly the significance of the paper.

It is a very good idea to keep all of these people informed about your publication as soon as it is published. You really want them to hear about it from you first and not from somebody else. This includes contacting any authorities that have issued permits. You may also want to contact funders.

>>>>

The day your paper is accepted, tell your co-authors. If they are in the same physical location as you then buy a cake, and celebrate with them at tea time. If cake isn’t your thing, then find another appropriate treat for you and your co-authors. It’s a great achievement and something you should share together. If you aren’t in the same location then make a plan for next time you meet together, even if that’s only online. There is real value in celebrating the positive times in your life as an Early Career Researcher. We all know that sharing the times when you are [rejected](#reject) will improve that experience. Similarly, celebrating the good times will give you more impetus to keep pushing through the down times. Extending this positive feeling to your co-authors, or those in your laboratory will always be appreciated.

Remember to make sure that all of your co-authors have a copy of the accepted version of the manuscript. Sometimes referred to as a **postprint**, you and your co-authors should submit this to your institutional repository so that the article can be reached as [Green Open Access](#green) by anyone who is interested in reading it. Another option if your library does not have a repository is to use an Open Source service like [Share Your Paper](https://shareyourpaper.org/). Note that when self archiving, you will need to use the publisher's [DOI](#DOI) in order to make your manuscript findable.

@@

## Take your time with proofs {#proofs}

The next step in the publishing process, once your paper has been accepted, is that it will go for type setting. Depending on the journal and the publisher, this process can proceed in several ways. Typically you will receive a notification that your paper proofs are ready and that you need to check and return them within 48 hours.

Some journals may have an additional __copy editing__ (also known as sub-editing) step. This is essentially an extra step checking journal style, syntax, punctuation, spelling and grammatical consistency. The copy editor may be tasked with checking additional issues, like species names (taxonomic authorities), names of chemical compounds used, etc. All will check citations against literature cited, and check that literature cited is sufficient to link with [CrossRef](#CrossRef). Most journals that I have worked with roll copy editing and proofing into a single step.

If you haven't checked proofs before, then it is important that you read the instructions from the publishers, that come with the proofs, carefully. They should tell you exactly what to do and if you are unsure about anything then talk to your co-authors.

Typically the publishers will send you a set of queries (copy edits) that relate to your proofs. They always ask you to check every author's name and affiliation. Other typical errors are that there are citations in the text that are not in the references. Or that there's literature in the references that are not cited.

The process of checking the proofs is very important. Errors can creep in during the type setting stage. Equally, copy editors can get it wrong, and you should read the text very carefully to make sure that errors have not been introduced. Pay special attention to the tables, table legends and figure legends.

You may also have the opportunity to change the size or orientation of figures, especially if it looks like they are not well presented in the proof. If the journal prints into columns, they may choose to put your figure in one column instead of across two. Another option is to have your figure in landscape across the whole page. Journals are generally pushed for space and so may refuse some requests for more room for larger figures. But you might get lucky if you make a good case.

Although the proofs are the responsibility of the corresponding author, it's good to get as many eyes on them as possible in order to spot any possible errors. Some journals let you know when proofs are likely to arrive, in which case it's a good idea to alert your co-authors and ask them whether they are prepared to look at them within that temporal window. I usually suggest that you make all your own corrections first before sending them around. Worst is for everyone to make their own corrections independently, as you'll have to put them all onto one set of proofs, and everyone is likely to spot some of the same errors.

Many publishers want to proofs back in a hurry (typically 48 hours). If you don't have the confidence to correct proofs yourself, or cannot pass it around your co-authors within this deadline, then you can simply write back and ask for an extended deadline for your proofs. It is important to get it right, and much better than having to correct the paper later.

@@

Probably the easiest way of doing proofs is to print them out and go through them with a pencil first. This allows you to take your time and you're more likely to spot errors this way than on the screen. However, some publishers will require you to submit proofs in an online system (effectively working with their LaTeX document). You should still have an opportunity to print and take your time with the proofs though, and these systems do allow for this.

Sometimes, there are a lot of problems with proofs, and you may not have the confidence that the publishers will make all the corrections as indicated. In this case, you should ask to see another round of proofs before committing. Most publishers will happily send you a second round of proofs.

@@

Sometimes your article may acquire several DOIs, for example from the publisher and from the preprint server. Sites like figshare and ResearchGate, assign different DOIs to content uploaded to their platforms. These are DOIs but they’re assigned by DataCite, another DOI registration agency. The different registration agencies provide different services that relate to DOIs and their associated metadata (from construction to the movie industry), and have different requirements for their member organisations. To prevent ambiguity, you should always use the DOI associated with the [Version of Record](#VoR). The biggest player registering content in scholarly output is [CrossRef](https://www.crossref.org/).

@@

Publishers normally deposit DOIs and other metadata (authors' names and addresses, publication dates, title, licence, funders, etc.) around the time the article is published online. This is called content registration. This is not just for articles, but also preprints, conference proceedings, books, book chapters, peer reviews and more can all have their content registered with CrossRef. The standardisation of this metadata means that not only is it possible to immediately find your article (the DOI is a link), but all of the metadata associated with it can be cross-referenced. This is of great help to funders, for example, who can look up all the products that their funding has produced without having to contact all of the people that they funded.

### DOI Tools
If you can't find the DOI for an article that you want to cite, then there is this very useful online software that will provide the DOI if there is one for every reference you enter: https://doi.crossref.org/simpleTextQuery or use the CrossRef Metadata Search: https://search.crossref.org/references

There’s an equally useful database that provides BibTex for DOIs that you enter: https://doi2bib.org/

## Add your paper to your ORCID account {#ORCID}
The ORCID (https://orcid.org/) number was devised to provide a common platform for authors to curate, and now some journals and funding authorities won’t allow you to submit or apply without one. This helps when there are authors with common (or even identical) names. As an NGO dedicated to helping scientists ascribe [credit for their work](#rescognito), it should be supported. ORCID will also help prevent author fraud (see [Part I](#authors)).

If you added your ORCID number to the metadata when you uploaded your manuscript, your ORCID record should be updated automatically (through Crossref) when the DOI of your article appears online. Inside your ORCID account, you can grant permission for Crossref to update the records automatically, otherwise you will need to log onto ORCID from time to time and approve changes suggested there. Alternatively, you can add works to your ORCID record via [Crossref Metadata Search](https://search.crossref.org/references).

Be careful not to create duplicate accounts, and make your ORCID available to your collaborators so that they aren't left guessing at your identity when it comes to submitting a paper. ORCID allows you to create some useful links for [your website](#website) or a QR code for adding to posters and presentations. If you find you have a duplicated account, it is simple to remove it (see [here](https://support.orcid.org/hc/en-us/articles/360006896634)).

## Once you have a publication date
Once your paper is published you have an opportunity to publicise it yourself. There are lots of different ways to do this, [see the next chapters](#populararticle). Some funders and institutions will want to know about press worthy publications before they are published, so that they can prepare a [press release](#pressrelease).

## On the day you publish
This is a great opportunity to contact all the people who helped you in your study and send them a PDF of your paper. The easiest way to do this is to go to the acknowledgements section and write an email that includes everyone that is mentioned in the acknowledgements. Write them a nice email in which you thank them for their help and explain briefly the significance of the paper.

It is a very good idea to keep all of these people informed about your publication as soon as it is published. You really want people in [your network](#network) to hear about it from you first and not from somebody else. This includes contacting any authorities that have issued permits. You may also want to contact funders.
